Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2021-04-18 manuscriptPGPEsuppl.tex
Fixed bunch of latex syntax errors.

\author{\scriptsize Am{\'e}lie Bacle}
\affiliation{Laboratoire Coopératif "Lipotoxicity and Channelopathies – ConicMeds", Université de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }

@@

\author{Patrick F.J. Fuchs}
\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
\alsoaffiliation{Université de Paris, UFR Sciences du Vivant, 75013, Paris, France}

@@

\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
A pure POPC system and a 50:50 POPC:POPE mixture were built and equilibrated using CHARMM-GUI [https://doi.org/10.1021/acs.jctc.5b00935]. They contained 256 lipids (for the mixture 64 lipids per leaflet for each species) and 34 water molecules per lipid. No ions were added. The production simulations were run for 300 ns with a time step of 2 fs. The first 50 ns were discarded for the analysis. The simulations were run with the GROMACS 2016.4 [https://doi.org/10.1016/j.softx.2015.06.001] version. The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 1 ps; lipids and water were coupled separately to the heat bath. Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 5.0 ps. A real space cut-off of 1.2 nm was employed for electrostatic interactions while the long-range part was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117]. A force-based switching function was used to switch the Lennard-Jones forces to zero over a range of 1−1.2 nm. All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H]. Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805]. The simulation files are available from ref 64 [https://doi.org/10.5281/zenodo.1306800] (pure POPC) and 65 [https://doi.org/10.5281/zenodo.1306821] (POPC:POPE mixture).

@@

Force field details
Simulations of POPC:POPE mixtures using the OPLS-MacRog force field [https://doi.org/10.1021/jp5016627] were performed. The initial force field parameter files in GROMACS format were taken from ref [https://doi.org/10.1016/j.dib.2016.03.067]. However, a number of errors were detected in the published files so we fixed them in the following way. For POPE, the two aliphatic chains sn-1 and sn-2 were switched (in the file the lipid was in fact OPPE); one atom (named C27) was not connected to the previous atom in the aliphatic chain so bonds, angles, dihedrals, pairs were included to create the connection; two atoms were called C27 thus one of them was renamed C28. For POPC, some impropers were missing leading to non planar systems for double bonds (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), we thus added back those impropers. The resulting fixed files can be found on Zenodo (https://zenodo.org/record/3741793/files/popc_fixed.itp, https://zenodo.org/record/3725637/files/pope_fixed.itp). To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results of ref [https://doi.org/10.1021/acs.jpcb.5b04878]. We found very similar values.
All the details of the procedure described here are described in https://www.dsimb.inserm.fr/~fuchs/project_Samuli/POPC_POPE/report_results_comparison.pdf.
The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].

>>>>

\author{\scriptsize Am{\'e}lie Bacle}
\affiliation{Laboratoire Coop\'{e}ratif "Lipotoxicity and Channelopathies – ConicMeds", Universit\'{e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }

@@

\author{Patrick F.J. Fuchs}
\affiliation{Sorbonne Universit\'{e}, Ecole Normale Sup\'{e}rieure, PSL University, CNRS, Laboratoire des Biomol\'{e}cules (LBM), 75005 Paris, France}
\alsoaffiliation{Universit\'{e} de Paris, UFR Sciences du Vivant, 75013, Paris, France}

@@

\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
A pure POPC system and a 50:50 POPC:POPE mixture were built and equilibrated using CHARMM-GUI [https://doi.org/10.1021/acs.jctc.5b00935].
They contained 256 lipids (for the mixture 64 lipids per leaflet for each species) and 34 water molecules per lipid.
No ions were added.
The production simulations were run for 300 ns with a time step of 2 fs.
The first 50 ns were discarded for the analysis.
The simulations were run with the GROMACS 2016.4 [https://doi.org/10.1016/j.softx.2015.06.001] version.
The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 1 ps; lipids and water were coupled separately to the heat bath.
Pressure was kept constant at 1 bar using a semi-isotropic Parrinello--Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 5.0 ps.
A real space cut-off of 1.2 nm was employed for electrostatic interactions while the long-range part was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117].
A force-based switching function was used to switch the Lennard-Jones forces to zero over a range of 1--1.2 nm.
All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
The simulation files are available from ref 64 [https://doi.org/10.5281/zenodo.1306800] (pure POPC) and 65 [https://doi.org/10.5281/zenodo.1306821] (POPC:POPE mixture).

@@

Force field details
Simulations of POPC:POPE mixtures using the OPLS-MacRog force field [https://doi.org/10.1021/jp5016627] were performed.
The initial force field parameter files in GROMACS format were taken from ref [https://doi.org/10.1016/j.dib.2016.03.067].
However, a number of errors were detected in the published files so we fixed them in the following way.
For POPE, the two aliphatic chains sn-1 and sn-2 were switched (in the file the lipid was in fact OPPE); one atom (named C27) was not connected to the previous atom in the aliphatic chain so bonds, angles, dihedrals, pairs were included to create the connection; two atoms were called C27 thus one of them was renamed C28.
For POPC, some impropers were missing leading to non planar systems for double bonds (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), we thus added back those impropers.
The resulting fixed files can be found on Zenodo {\url{https://zenodo.org/record/3741793/files/popc_fixed.itp}, \url{https://zenodo.org/record/3725637/files/pope_fixed.itp}).
To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results of ref [https://doi.org/10.1021/acs.jpcb.5b04878]. We found very similar values.
All the details of the procedure described here are described in \url{https://www.dsimb.inserm.fr/~fuchs/project_Samuli/POPC_POPE/report_results_comparison.pdf}.
The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].
